---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Setup

```{r}
library(ggplot2)
```

```{r, cache=T}
data <- read.csv('activity.csv')
```

## What is the mean total number of steps taken per day?

```{r totalNumberOfSteps}
stepsPerDay <- aggregate(list(Steps=data$steps), by = list(Date=data$date), sum, na.rm = T)
ggplot(data=stepsPerDay, mapping = aes(Steps)) + geom_histogram(binwidth=706)
```

mean and median steps per day:

```{r mediaAndMean, cache=T}
median(stepsPerDay$Steps)
mean(stepsPerDay$Steps)
```

## What is the average daily activity pattern?

Average number of steps per interval:

```{r stepsByInterval, cache=T}
avgStepsByInterval <- aggregate(list(Steps=data$steps), by = list(Interval=data$interval), mean, na.rm = T)
ggplot(data=avgStepsByInterval, mapping = aes(x=Interval, y=Steps)) + geom_line()
```

max number of steps:

```{r}
max(avgStepsByInterval$Steps)
```

## Imputing missing values

Total number of rows with missing values:

```{r}
nrow(data[!complete.cases(data),])
```

Set `NA`s to the mean of each interval (since this was one possible acceptable choice offered by the assignment) and create a new histogram

```{r, cache=T}
data2 <- data
for (x in 1:nrow(data2)){
    vec <- data2[x,]
    if(is.na(vec$steps)){
        data2[x,]$steps <- avgStepsByInterval[avgStepsByInterval$Interval == vec$interval,]$Steps
    }
}
stepsPerDay <- aggregate(list(Steps=data2$steps), list(Date=data2$date), sum, na.rm = T)
ggplot(data=stepsPerDay, mapping = aes(Steps)) + geom_histogram(binwidth=706)
```

New median and mean

mean and median steps per day:

```{r}
median(stepsPerDay$Steps)
mean(stepsPerDay$Steps)
```

## Are there differences in activity patterns between weekdays and weekends?

Create a new column of type `Factor` with levels `weekend` and `weekday`:

```{r}
Sys.setlocale("LC_TIME", "en_US")
weakdays <- weekdays(strptime(data$date,'%Y-%m-%d'))
weekendify <- function(weekday){
    if (weekday == "Saturday" || weekday == "Sunday"){
        "weekend"
    } else {
        "weekday"
    }
}
data$isWeekend_p <- factor(sapply(weakdays, weekendify))
```

Plot activity patterns for weekends vs. weekdays:

```{r}
avgStepsByInterval <- aggregate(list(Steps=data$steps), list(Interval=data$interval, data$isWeekend_p), mean, na.rm = T)
g <- ggplot(data=avgStepsByInterval, mapping = aes(x=Interval, y=Steps)) + geom_line()
g + facet_grid(Group.2 ~ .)
```
